---
title: "R base functions for mass_dataset"
author:
- name: Xiaotao Shen (https://www.shenxt.info/)
date: "Created on 2021-12-04 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{base_function}
  %\VignettePackage{massdataset}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

`mass_dataset` object support many R base functions.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(massdataset)
library(tidyverse)

data("expression_data")
data("sample_info")
data("sample_info_note")
data("variable_info")
data("variable_info_note")

object =
  create_mass_dataset(
    expression_data = expression_data,
    sample_info = sample_info,
    variable_info = variable_info,
    sample_info_note = sample_info_note,
    variable_info_note = variable_info_note
  )
```

For example, you can get the information of your object.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
dim(object)
nrow(object)
ncol(object)
dimnames(object)
```

This means that `object` has 1000 variables and 8 samples.

You can also get the sample ids and variables.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
colnames(object)
head(rownames(object))
```


Use `[` to select variables and samples from object.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
##only remain first 5 variables
object[1:5,]

##only remain first 5 samples
object[,1:5]

##only remain first 5 samples and 5 variables
object[1:5,1:5]
```

If you know the variables or sample names you want to select, you can also use the samples ids or variables ids.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
colnames(object)
object[,c("Blank_3", "Blank_4")]

```


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
###log
object2 = 
  log(object + 1, 10)
unlist(object[1,,drop = TRUE])
unlist(object2[1,,drop = TRUE])

###scale
object2 = 
  scale(object, center = TRUE, scale = TRUE)
unlist(object[1,,drop = TRUE])
unlist(object2[1,,drop = TRUE])

```

# **Session information**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```