---
title: "Add MS2 spectra data into mass_dataset class object"
author:
- name: Xiaotao Shen
date: "Created on 2021-12-04 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{Add MS2 spectra data into mass_dataset class object}
  %\VignettePackage{massdataset}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

`mass_data` class object can also contain MS2 data. 

# **Data preparation**

## mass_data class object

We need to create a `mass_data` class object first; see [this document](https://www.tidymass.org/massdataset/articles/data_import_and_export.html). Here we use the demo datasets bundled with the package.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=FALSE}
library(massdataset)

data("expression_data")
data("sample_info")
data("variable_info")

object_pos <-
  create_mass_dataset(
    expression_data = expression_data,
    sample_info = sample_info,
    variable_info = variable_info
  )
```

## MS2 data

The MS2 raw data should be converted to mgf format data. Please refer to [this document](https://www.tidymass.org/metid/articles/database_construction.html).

Here we use the MGF example bundled with `massdataset`.

# **Add MS2 to `mass_dataset` class object**

Positive mode.

```{r,eval=FALSE,warning=FALSE, R.options="", message=FALSE, cache=FALSE}
ms2_dir <- system.file("ms2_data", package = "massdataset")

object_pos2 =
  mutate_ms2(
    object = object_pos,
    column = "rp",
    polarity = "positive",
    ms1.ms2.match.mz.tol = 10,
    ms1.ms2.match.rt.tol = 15,
    path = ms2_dir
  )
```

# **Session information**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=FALSE}
sessionInfo()
```
