---
title: "Mutate MS2 into mass_dataset"
author:
- name: Xiaotao Shen (https://www.shenxt.info/)
date: "Created on 2021-12-04 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{mutate_ms2}
  %\VignettePackage{massdataset}
  % \VignetteEngine{knitr::rmarkdown}
  % \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
```

`mass_data` class object can also contain MS2 data. 

# **Data preparation**

## **MS1 data**

We need to create a `mass_data` class object first, [see this document](https://tidymass.github.io/massdataset/articles/data_import_and_export.html). And here we use the data from this step as examples.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
load("feature_table/object_pos")
load("feature_table/object_neg")
```

## **MS2 data**

The MS2 raw data should be converted to mgf format data. Please [refer this document](https://tidymass.github.io/metid/articles/database_construction.html).

Here we use the demo data for `tidymass`, please download it and put it in the `feature_table` folder.

[Download linke is here.](https://drive.google.com/file/d/1lDf9i49jQQ-Ar98EMdj0EwhlmtDBbG0g/view?usp=sharing)

Then uncompress it.

# **Mutate MS2 to `mass_dataset` class object**

Positive mode.

```{r,eval=FALSE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
object_pos2 = 
mutate_ms2(
  object = object_pos,
  column = "rp",
  polarity = "positive",
  ms1.ms2.match.mz.tol = 10, 
  ms1.ms2.match.rt.tol = 15, 
  path = "feature_table/MS2_data/POS/"
)
```

Negative mode.

```{r,eval=FALSE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
object_neg2 = 
mutate_ms2(
  object = object_neg,
  column = "rp",
  polarity = "negative",
  ms1.ms2.match.mz.tol = 10, 
  ms1.ms2.match.rt.tol = 15, 
  path = "feature_table/MS2_data/NEG/"
)
```

# **Session information**

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```
